---
title: "XGBoost for Instacart"
author: "Boriss"
date: "July 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

cat("\014")  # clear console
rm(list=ls(all=TRUE))

require("knitr")
opts_knit$set(root.dir = 'c:/BBB/PROPACAD/CAPSTONE/BASKET/')


library(data.table)
library(dplyr)
library(tidyr)
```

```{r}
path <- "./DATA"

orders      <- fread(file.path(path,'orders.csv'               ) )
products    <- fread(file.path(path,'products.csv'             ) )
ordert      <- fread(file.path(path,'order_products__train.csv') )
orderp      <- fread(file.path(path,'order_products__prior.csv') )
aisles      <- fread(file.path(path,'aisles.csv'               ) )
departments <- fread(file.path(path,'departments.csv'          ) )
```

```{r}
# Reshape data ------------------------------------------------------------
aisles$aisle           <- as.factor(aisles$aisle)
departments$department <- as.factor(departments$department)
orders$eval_set        <- as.factor(orders$eval_set)
products$product_name  <- as.factor(products$product_name)
```

Substitute aisle_id and department_id with their names from aisles and department
Dataframe products before
```{r}
head( products )
```
and after

```{r}
products <- products %>% 
  inner_join(aisles) %>% inner_join(departments) %>% 
  select(-aisle_id, -department_id)
rm(aisles, departments) # remove (aisles, departments) from memory
head( products)
```

### Add 
```{r}
head( orders )
```
```{r}
head( ordert, n = 10 )
```

**match** returns a vector of the positions of (first) matches of its first argument in its second.

```{r}
ordert$order_id[1:10]
#NB! orders contain order_id's in prior, train, test 
#match(ordert$order_id[1:10], orders$order_id) returns indexes of order_id in ordert in orders

orders[ match(ordert$order_id[1:10], orders$order_id), ]

```

### Add corresponding user_id to ordert
```{r}
ordert$user_id <- orders$user_id[ match(ordert$order_id, orders$order_id) ]
head( ordert, n = 10 )
```

```{r}
head( orderp, n = 10 )
head( orders, n = 10 )
```
### Expand orders with detailed information per order from orderp
```{r}
orders_products <- orders %>% inner_join(orderp, by = "order_id")
head(orders_products, n = 25 )
unique( orders_products$eval_set ) #check whether orders are only from prior
rm(orderp)
gc()
```

### Check how many unique products were ordered in prior 
```{r}
hist( orders_products$product_id, breaks = 1000 )
length( unique( orders_products$product_id ) )
dim( products)
```


arrange {dplyr}: arrange(user_id, order_number, product_id) sorts orders_products in ascending order 
- first according to user_id
- second according to order_number
- third accorindg to product_id

mutate(product_time = row_number()): i.e. row_number(): equivalent to rank(ties.method = "first")
create a variable that tracks how many times each product_id was ordered by each user_id
```{r}
prd_tmp <- orders_products %>%
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup()
head( prd_tmp, n = 40 )
```
```{r}
prd_tmp2 = head( prd_tmp, n = 140 )
prd_tmp %>% filter( product_id == 196 ) %>% select(user_id, product_id, product_time) %>% head( n = 30 )
```

### Determine product characteristics:
-- prod_orders: how many times a product_id was ordered by all users
-- prod_reorders: how many times a product_id was reordered by all users
-- prod_first_orders: number of times a product_id was ordered for the first time
-- prod_second_order: number of times a product_id was ordered for the second time
```{r} 

# Products ----------------------------------------------------------------
rm( prd_tmp, prd_tmp2 )

prd <- orders_products %>%
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup() %>%
  group_by(product_id) %>%
  summarise(
    prod_orders        = n(),
    prod_reorders      = sum(reordered),
    prod_first_orders  = sum(product_time == 1),
    prod_second_orders = sum(product_time == 2)
  )
prd_tmp3 = head( prd, n = 40 )
```

### create several features
prod_reorder_probability - probability of first re-order
prod_reorder_times       - aka rd$prod_orders / prd$prod_first_orders as 
                           prd$prod_first_orders + prd$prod_reorders = aka rd$prod_orders see prd_tmp3
                         : inverse of share of prod_first_orders in prod_orders, if that share is close to one,
                           then it is very likely that this product on average was bought only once
prod_reorder_ratio       - share of reorders in orders

```{r}
prd$prod_reorder_probability <- prd$prod_second_orders / prd$prod_first_orders
prd$prod_reorder_times       <- 1 + prd$prod_reorders  / prd$prod_first_orders
prd$prod_reorder_ratio       <- prd$prod_reorders      / prd$prod_orders

prd <- prd %>% select(-prod_reorders, -prod_first_orders, -prod_second_orders)

rm(products)
gc()

prd_tmp4 = head( prd, n = 40 )

hist( prd$prod_reorder_ratio, breaks = 1000)

```

```{r}
hist( prd$prod_reorder_ratio, breaks = 5000)
sum( prd$prod_reorder_ratio == 0) 

barplot( cumsum( prd$prod_reorder_ratio[ order( prd$prod_reorder_ratio ) ] ),
         main = "Ordered reorder ratio by product_id" )

summary( prd$prod_reorder_ratio )
```

```{r}

# Products ----------------------------------------------------------------
prd <- orders_products %>%
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup() %>%
  group_by(product_id) %>%
  summarise(
    prod_orders = n(),
    prod_reorders = sum(reordered),
    prod_first_orders = sum(product_time == 1),
    prod_second_orders = sum(product_time == 2)
  )

prd$prod_reorder_probability <- prd$prod_second_orders / prd$prod_first_orders
prd$prod_reorder_times <- 1 + prd$prod_reorders / prd$prod_first_orders
prd$prod_reorder_ratio <- prd$prod_reorders / prd$prod_orders

prd <- prd %>% select(-prod_reorders, -prod_first_orders, -prod_second_orders)

rm(products)
gc()

```

### characteristics of users in prior set
- user_order_number         : order_number per user
- user_period               : how long a user_id was a customer
- user_mean_days_since_prior: average shopping frequency
```{r}
# Users -------------------------------------------------------------------
users <- orders %>%
  filter(eval_set == "prior") %>%
  group_by(user_id) %>%
  summarise(
    user_orders = max(order_number),
    user_period = sum(days_since_prior_order, na.rm = T),
    user_mean_days_since_prior = mean(days_since_prior_order, na.rm = T)
  )
users_tmp = head( users, n = 40 )
```