---
title: "XGBoost for Instacart"
author: "Boriss"
date: "July 10, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

cat("\014")  # clear console
rm(list=ls(all=TRUE))

require("knitr")
#opts_knit$set(root.dir = 'c:/BBB/PROPACAD/CAPSTONE/BASKET/')      #T520
#opts_knit$set(root.dir = 'c:/Users/sboriss/BBB/PROPACAD/')         # KOF
#opts_knit$set(root.dir = 'U:/BBB/PROPACAD/')                        # MTEC-0373

library(data.table)
library(dplyr)
library(tidyr)
```

```{r}
path_main = "U:/BBB/PROPACAD/"
path_data = paste0(path_main, 'DATA/')
path_subm = paste0(path_main, 'SUBM/')

orders      <- fread( paste0(path_data, 'orders.csv') ) 
products    <- fread( paste0(path_data, 'products.csv'             ) )
ordert      <- fread( paste0(path_data, 'order_products__train.csv') )
orderp      <- fread( paste0(path_data, 'order_products__prior.csv') )
aisles      <- fread( paste0(path_data, 'aisles.csv'               ) )
departments <- fread( paste0(path_data, 'departments.csv'          ) )

ordert_tmp = head( ordert, n = 20 )

```


```{r}
#path <- "./DATA"

#orders      <- fread(file.path(path,'orders.csv'               ) )
#products    <- fread(file.path(path,'products.csv'             ) )
#ordert      <- fread(file.path(path,'order_products__train.csv') )
#orderp      <- fread(file.path(path,'order_products__prior.csv') )
#aisles      <- fread(file.path(path,'aisles.csv'               ) )
#departments <- fread(file.path(path,'departments.csv'          ) )
```

```{r}
# Reshape data ------------------------------------------------------------
aisles$aisle           <- as.factor(aisles$aisle)
departments$department <- as.factor(departments$department)
orders$eval_set        <- as.factor(orders$eval_set)
products$product_name  <- as.factor(products$product_name)
```

Substitute aisle_id and department_id with their names from aisles and department
Dataframe products before
```{r}
head( products )
```
and after

```{r}
products <- products %>% 
  inner_join(aisles) %>% inner_join(departments) %>% 
  select(-aisle_id, -department_id)
rm(aisles, departments) # remove (aisles, departments) from memory
head( products)
```

### Add 
```{r}
head( orders )
```
```{r}
head( ordert, n = 10 )
```

**match** returns a vector of the positions of (first) matches of its first argument in its second.

```{r}
ordert$order_id[1:10]
#NB! orders contain order_id's in prior, train, test 
#match(ordert$order_id[1:10], orders$order_id) returns indexes of order_id in ordert in orders

orders[ match(ordert$order_id[1:10], orders$order_id), ]

```

### Add corresponding user_id to ordert
```{r}
ordert$user_id <- orders$user_id[ match(ordert$order_id, orders$order_id) ]
ordert_tmp1= head( ordert, n = 10 )
```

```{r}
head( orderp, n = 10 )
head( orders, n = 10 )
```
### Expand orders with detailed information per order from orderp
```{r}
orders_products     = orders %>% inner_join(orderp, by = "order_id")
orders_products_tmp = head(orders_products, n = 250 )
unique( orders_products$eval_set ) #check whether orders are only from prior
rm(orderp)
gc()
```

### Check how many unique products were ordered in prior 
```{r}
hist( orders_products$product_id, breaks = 1000 )
length( unique( orders_products$product_id ) )
dim( products)
```


arrange {dplyr}: arrange(user_id, order_number, product_id) sorts orders_products in ascending order 
- first according to user_id
- second according to order_number
- third accorindg to product_id

mutate(product_time = row_number()): i.e. row_number(): equivalent to rank(ties.method = "first")
create a variable that tracks how many times each product_id was ordered by each user_id
```{r}
prd_tmp <- orders_products %>%
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup()
head( prd_tmp, n = 40 )
```
```{r}
prd_tmp2 = head( prd_tmp, n = 140 )
prd_tmp %>% filter( product_id == 196 ) %>% select(user_id, product_id, product_time) %>% head( n = 30 )
```

### Determine product characteristics:
- prod_orders: how many times a product_id was ordered by all users
- prod_reorders: how many times a product_id was reordered by all users
- prod_first_orders: number of times a product_id was ordered for the first time
- prod_second_order: number of times a product_id was ordered for the second time
```{r} 

# Products ----------------------------------------------------------------
rm( prd_tmp, prd_tmp2 )

prd <- orders_products %>%
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup() %>%
  group_by(product_id) %>%
  summarise(
    prod_orders        = n(),
    prod_reorders      = sum(reordered),
    prod_first_orders  = sum(product_time == 1),
    prod_second_orders = sum(product_time == 2)
  )
prd_tmp3 = head( prd, n = 40 )
```

### create several features
prod_reorder_probability - probability of first re-order
prod_reorder_times       - aka prd\$prod\_orders / prd\$prod\_first\_orders as 
                           prd\$prod_first\_orders + prd\$prod\_reorders = aka rd\$prod\_orders see prd\_tmp3
                         : inverse of share of prod_first_orders in prod_orders, if that share is close to one,
                           then it is very likely that this product on average was bought only once
prod_reorder_ratio       - share of reorders in orders

```{r}
prd$prod_reorder_probability <- prd$prod_second_orders / prd$prod_first_orders
prd$prod_reorder_times       <- 1 + prd$prod_reorders  / prd$prod_first_orders
prd$prod_reorder_ratio       <- prd$prod_reorders      / prd$prod_orders

prd <- prd %>% select(-prod_reorders, -prod_first_orders, -prod_second_orders)

rm(products)
gc()

prd_tmp4 = head( prd, n = 40 )

hist( prd$prod_reorder_ratio, breaks = 1000)

```

```{r}
hist( prd$prod_reorder_ratio, breaks = 5000)
sum( prd$prod_reorder_ratio == 0) 

barplot( prd$prod_reorder_ratio[ order( prd$prod_reorder_ratio ) ],
         main = "Ordered reorder ratio by product_id" )

summary( prd$prod_reorder_ratio )
```

```{r}

# Products ----------------------------------------------------------------
prd <- orders_products %>%
  arrange(user_id, order_number, product_id) %>%
  group_by(user_id, product_id) %>%
  mutate(product_time = row_number()) %>%
  ungroup() %>%
  group_by(product_id) %>%
  summarise(
    prod_orders = n(),
    prod_reorders = sum(reordered),
    prod_first_orders = sum(product_time == 1),
    prod_second_orders = sum(product_time == 2)
  )

prd$prod_reorder_probability <- prd$prod_second_orders / prd$prod_first_orders
prd$prod_reorder_times <- 1 + prd$prod_reorders / prd$prod_first_orders
prd$prod_reorder_ratio <- prd$prod_reorders / prd$prod_orders

prd <- prd %>% select(-prod_reorders, -prod_first_orders, -prod_second_orders)

rm(products)
gc()

```

## characteristics of users in prior set
- user_order_number         : order_number per user
- user_period               : how long a user_id was a customer
- user_mean_days_since_prior: average shopping frequency
```{r}
# Users -------------------------------------------------------------------
users <- orders %>%
  filter(eval_set == "prior") %>%
  group_by(user_id) %>%
  summarise(
    user_orders = max(order_number),
    user_period = sum(days_since_prior_order, na.rm = T),
    user_mean_days_since_prior = mean(days_since_prior_order, na.rm = T)
  )
users_tmp = head( users, n = 40 )
```

- user_total_products: total number of products ordered
- user_reorder_ratio: share of products reordered in orders_id[2:last]; 
  there were no reorders in the first order [order_number = 1]
- user_distinct_products: number of unique products ordered
```{r}
us <- orders_products %>%
  group_by(user_id) %>%
  summarise(
    user_total_products = n(),                                        #how many products were purchased in total
    user_reorder_ratio = sum(reordered == 1) / sum(order_number > 1), #NB! why divide by sum(order_number > 1)?
    user_distinct_products = n_distinct(product_id)                   #how many distict products were purchased
  )
```

### Add columns ( user_total_products, user_reorder_ratio, user_distinct_products)  from us to user
- user_average_basket - size of average basket per order
```{r}
users <- users %>% inner_join(us)
users$user_average_basket <- users$user_total_products / users$user_orders

```
```{r}
users_tmp1 = head( users, n = 40 )
```

```{r}
plot( users$user_average_basket, users$user_reorder_ratio)
```

#add order_id from train/test

```{r}
orders_tmp = head( orders, n = 50)
```

```{r}
us <- orders %>%
  filter(eval_set != "prior") %>%
  select(user_id, order_id, eval_set,
         days_since_last_order = days_since_prior_order
         )

users <- users %>% inner_join(us)

rm(us)
gc()

```

```{r}
users_tmp2 = head( users, n = 50)
```

# Database ----------------------------------------------------------------
### create features
- up\_orders: how many times a product\_id was ordered by user\_id
- up_first_order: in which order a product_id was first ordered
- up_last_order:  in which order a product_id was last  ordered

```{r}
data <- orders_products %>%
  group_by(user_id, product_id) %>% 
  summarise(
    up_orders = n(),
    up_first_order = min(order_number),
    up_last_order  = max(order_number),
    up_average_cart_position = mean(add_to_cart_order)
  )

data_tmp = head( data, n = 250 )


rm(orders_products, orders)
```

```{r}
data <- data %>% 
  inner_join(prd, by = "product_id") %>%
  inner_join(users, by = "user_id")

data_tmp1 = head( data, n = 250 )
```

### create features
- up\_order\_rate: user-specific order intensity of a product_id
                 : proportion of times a product_id was ordered in total number of orders
- up\_orders\_since\_last_order: how many orders ago a product was last ordered
- up_order_rate_since_first_order: product-specific order intensity over the time between a product was first ordered and the last order
```{r}
data$up_order_rate                   <- data$up_orders / data$user_orders
data$up_orders_since_last_order      <- data$user_orders - data$up_last_order
data$up_order_rate_since_first_order <- data$up_orders / (data$user_orders - data$up_first_order + 1)

data_tmp2 = head( data, n = 250 )
```


#add train 
```{r}
data <- data %>% 
  left_join(ordert %>% select(user_id, product_id, reordered), 
            by = c("user_id", "product_id"))

data_tmp3 = head( data, n = 250 )

rm(ordert, prd, users)
gc()
```

##there are no labels 'prior' in data
```{r}
data_prior = data %>% filter( eval_set == "prior")
data_train = data %>% filter( eval_set == "train")
data_test  = data %>% filter( eval_set == "test")

dim( data_prior )
dim( data_train )
dim( data_test )

#ordert_user_id_1 = ordert %>% filter( user_id == 1 ) %>% arrange( product_id )
```


# Train / Test datasets ---------------------------------------------------
```{r}
train <- as.data.frame(data[data$eval_set == "train",])
train$eval_set <- NULL
train$user_id <- NULL
train$product_id <- NULL
train$order_id <- NULL
train$reordered[is.na(train$reordered)] <- 0

train_tmp = head( train, n = 250 )

test <- as.data.frame(data[data$eval_set == "test",])
test$eval_set <- NULL
test$user_id <- NULL
test$reordered <- NULL
#NB! keep order_id and product_id: use later to predict the basket

test_tmp = head( test, n = 250 )

rm(data)
gc()
```

```{r}
dim( train )
colnames( train )
dim( test )
colnames( test )
```

```{r}
library(xgboost)
library(Ckmeans.1d.dp)

params <- list(
  "objective"           = "reg:logistic",
  "eval_metric"         = "logloss", # rmse
  "eta"                 = 0.2,
  "max_depth"           = 5,
  "min_child_weight"    = 1,  #10
  "gamma"               = 0.70,
  "subsample"           = 0.77,
  "colsample_bytree"    = 0.95,
  "alpha"               = 2e-05,
  "lambda"              = 6 #10
)

subtrain <- train %>% sample_frac(0.15)
subtrain_tmp = head(subtrain, n = 100) 

ch_stdn_data <- FALSE

if( ch_stdn_data ){
  
  # standardize features in subtrain set
  subtrain_mat      = as.matrix(subtrain %>% select(-reordered) )
  subtrain_mat_stnd = scale( subtrain_mat ); rm( subtrain_mat )
  X                 = xgb.DMatrix(subtrain_mat_stnd, label = subtrain$reordered ); rm( subtrain_mat_stnd )  
  
}else{
  
  X = xgb.DMatrix( as.matrix(subtrain %>% select(-reordered) ), label = subtrain$reordered ); 
  
}


if(FALSE){
  cv <- xgb.cv(data = X, params = params, nrounds =100, nthread = 8, nfold = 5)
}
```

```{r}
model = xgboost(data = X, params = params, nrounds = 100, nthread = 8)

importance <- xgb.importance(colnames(X), model = model)
xgb.ggplot.importance(importance)

rm(X, importance, subtrain)
gc()
```

# Predict with model -------------------------------------------------------------
```{r}
#rm( train )
if( ch_stdn_data ){
  
   X <- xgb.DMatrix( scale( as.matrix(test %>% select(-order_id, -product_id)) ) )
  
}else{
  
   X <- xgb.DMatrix( as.matrix(test %>% select(-order_id, -product_id) ) ) 
   
}


test$reordered <- predict(model, X)

test_tmp1 = head( test, n = 250 )
```

```{r}
cutoff = 0.21

hist( test$reordered )
abline( v = cutoff , col = 2, lwd = 2)
mean( test$reordered > cutoff )
```

##compute average predicted probability of reordered
```{r}
test_1st <- test %>%
            group_by( order_id ) %>% 
            summarise(
            #pred_prob_reorder = median(   reordered)              #08
            #pred_prob_reorder = quantile( reordered, 0.75)        #09
            #pred_prob_reorder = quantile( reordered, 0.85)        #10
            #pred_prob_reorder = quantile( reordered, 0.75) + 0.1  #11
             pred_prob_reorder = quantile( reordered, 0.5) + 0.15  #12 LB 0.369
            
     )
test_1st_tmp = head( test_1st, n = 250 )
```

##put back order-specific mean of predicted probabilities of reorder
```{r}
test_2nd <- test_1st %>% 
            left_join( test, by = "order_id")

test_2nd_tmp = head( test_2nd, n = 250 )

test_2nd$reordered_in = ( test_2nd$reordered > test_2nd$pred_prob_reorder) * 1
#test_2nd$reordered_in = ( test_2nd$reordered > cutoff) * 1

test_2nd_tmp1 = head( test_2nd, n = 250 )

#compute predicted average basket size
pred_basket <- test_2nd %>%
              group_by( order_id ) %>%
              summarise(
                pred_basket_size = sum( reordered_in )
              )
pred_basket_tmp = head( pred_basket, n = 20 )              

summary( pred_basket_tmp$pred_basket_size )

submission <- test_2nd %>%
  filter(reordered_in == 1) %>%
  group_by(order_id) %>%
  summarise(
    products = paste(product_id, collapse = " ")
  )

submission_tmp = head( submission, n = 250 )

missing <- data.frame(
  order_id = unique(test$order_id[!test$order_id %in% submission$order_id]),
  products = "None"
)

submission <- submission %>% bind_rows(missing) %>% arrange(order_id)

submission_tmp = head( submission, n = 250 )

write.csv(submission, file = paste0( path_subm,"submit12_pred.csv"), row.names = F)
```

