---
title: "Instacart: Exploratory Analysis"
output: html_notebook
---

This part is based on the code provided at https://www.kaggle.com/philippsp/exploratory-analysis-instacart by Philipp Spachtholz. 

```{r "setup", include=FALSE}

cat("\014")  # clear console
rm(list=ls(all=TRUE))

require("knitr")
opts_knit$set(root.dir = 'c:/BBB/PROPACAD/CAPSTONE/BASKET/')
```


```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)

orders <- fread('./DATA/orders.csv')
products <- fread('./DATA/products.csv')
order_products_train <- fread('./DATA/order_products__train.csv')
order_products_prior <- fread('./DATA/order_products__prior.csv')
aisles <- fread('./DATA/aisles.csv')
departments <- fread('./DATA/departments.csv')
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

```{r}
head(orders)
```

```{r}
head(products)
```


```{r}
head(order_products_train, n = 20)
```
```{r}
head(aisles, n = 20)
```

```{r}
head(departments, n = 20)
```
