---
title: "Instacart: Exploratory Analysis"
output: html_notebook
---

This part is based on the code provided at https://www.kaggle.com/philippsp/exploratory-analysis-instacart by Philipp Spachtholz. 

```{r "setup", include=FALSE}

cat("\014")  # clear console
rm(list=ls(all=TRUE))

require("knitr")
opts_knit$set(root.dir = 'c:/BBB/PROPACAD/CAPSTONE/BASKET/')
```


```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library( mvoutlier )
```

```{r}
orders <- fread('./DATA/orders.csv')
products <- fread('./DATA/products.csv')
order_products_train <- fread('./DATA/order_products__train.csv')
order_products_prior <- fread('./DATA/order_products__prior.csv')
aisles <- fread('./DATA/aisles.csv')
departments <- fread('./DATA/departments.csv')
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

```{r}
head(orders)
```

```{r}
head(products)
dim(products)
```


```{r}
head(order_products_train, n = 20)
dim( order_products_train )
```


```{r}
head(order_products_prior, n = 20)
dim(order_products_prior)
```

```{r}
head(aisles, n = 20)
```

```{r}
head(departments, n = 20)
```

```{r}
orders
```


```{r}

```


Recode variables


```{r}
orders      <- orders      %>% mutate(order_hour_of_day = as.numeric(order_hour_of_day), eval_set = as.factor(eval_set))
products    <- products    %>% mutate(product_name      = as.factor(product_name))
aisles      <- aisles      %>% mutate(aisle             = as.factor(aisle))
departments <- departments %>% mutate(department        = as.factor(department))
```


Get user id's for train and test parts
```{r}
user_id_train = orders %>% filter(eval_set == "train") %>% 
                           select( user_id )
user_id_train = user_id_train[['user_id']]

length(user_id_train)

user_id_test = orders %>% filter(eval_set == "test") %>% 
                           select( user_id )
user_id_test = user_id_test[['user_id']]
length(user_id_test)
```

Create for each customer a density matrix. For those who repeatedly order the same products this matrix will be dense, and for
those who order different products each time this matrix will be sparse.

```{r}

ptm <- proc.time()

big_features <- lapply( user_id_train[1:300], function( user_id_i ){ # user_id_i = 5
  
    user_id_i_prior <- orders %>% filter(user_id==user_id_i, eval_set == 'prior' ) 
    user_id_i_prior <- user_id_i_prior %>% mutate( order_id_num = seq( nrow(user_id_i_prior) ) ) %>% 
                                           select( order_id,order_id_num, user_id )  %>% 
                                           left_join(order_products_prior,by="order_id") %>%
                                           left_join(products            , by = 'product_id')

    user_id_i_prior <- user_id_i_prior %>% mutate( product_id = as.factor(product_id) )
    user_id_i_prior
    
    user_id_i_prior_num_order   = max(            user_id_i_prior$order_id_num )
    user_id_i_prior_num_product = length( levels( user_id_i_prior$product_id   ) )
    
    #create an empty table
    user_id_i_prior_tbl = matrix( 0, nrow = user_id_i_prior_num_product, ncol = user_id_i_prior_num_order)
    rownames(user_id_i_prior_tbl) = unique( user_id_i_prior$product_name )
    colnames(user_id_i_prior_tbl) = levels(user_id_i_prior$order_id_num  )
    
    #insert values in the table
    for( i in seq( user_id_i_prior_num_order ) ){
    
      product_name   = user_id_i_prior %>% filter( order_id_num == i )
      product_name
      
      product_name   = product_name %>% select(product_name)
      product_name
      
      #convert from factor back to character
      product_name %>% mutate_if(is.factor, as.character) -> product_name
      product_name
      
      row_indx = as.vector( sapply( product_name[,1], function(x) which( x ==     rownames(user_id_i_prior_tbl) ) ) )
      row_indx
      
      user_id_i_prior_tbl[row_indx,i] = 1
    }
    
    num_purchases_in_prior = sum( colSums( user_id_i_prior_tbl ) )
    
    dens_tbl_i = num_purchases_in_prior / ( user_id_i_prior_num_order * user_id_i_prior_num_product)
    
    #get average order size in prior
    ave_size_order_prior_i = as.vector( colSums( user_id_i_prior_tbl ) ) %>% mean
    ave_size_order_prior_i
    
    #get the size of last order in prior
    lst_size_order_prior_i = as.vector( tail( colSums( user_id_i_prior_tbl), n = 1) ) 
    lst_size_order_prior_i
    
    #get average proportion of ordered items in second, third,..., last orders
    prop_reordered_ave_prior = user_id_i_prior %>% filter( order_id_num != 1) %>%
                                                   select( reordered ) %>% colMeans
    prop_reordered_ave_prior
    
    #get proportion of reordered items in the last order in
    prop_reordered_lst_prior = user_id_i_prior %>% filter( order_id_num == user_id_i_prior_num_order) %>%
                                                   select( reordered ) %>% colMeans                       
    prop_reordered_lst_prior           
    
    ########### extract from train ##############
    user_id_i_train <- orders %>% filter(user_id==user_id_i, eval_set == 'train')  %>% 
                                  select( order_id, user_id )  %>% 
                                  left_join(order_products_train,by="order_id") %>%
                                  left_join(products, by = 'product_id')
    user_id_i_train
    
    size_order_train     = nrow( user_id_i_train )
    prop_reordered_train = user_id_i_train %>% select( reordered )  %>% colMeans
    prop_reordered_train
    ########### extract from train ##############
    
    
    print( paste( "number of orders in prior: "  , user_id_i_prior_num_order   ) )
    print( paste( "number of products in prior: ", user_id_i_prior_num_product ) )
    
    
    #put features together
    df_features = data.frame( user_id                     = user_id_i, 
                              num_orders_in_prior         = user_id_i_prior_num_order,
                              num_products_in_prior       = user_id_i_prior_num_product,
                              num_purchases_in_prior      = num_purchases_in_prior,
                              density_order_products      = dens_tbl_i,
                              ave_ordersize_in_prior      = ave_size_order_prior_i,
                              lst_ordersize_in_prior      = lst_size_order_prior_i,
                              ave_prop_reordered_in_prior = prop_reordered_ave_prior,
                              lst_prop_reordered_in_prior = prop_reordered_lst_prior,
                              ordersize_in_train          = size_order_train,
                              prop_reordered_in_train     = prop_reordered_train )
    rownames(df_features) = user_id_i
    
    df_features
})
cat('\\n***EXECUTION TIME***:\n')
print( proc.time() - ptm )


big_features = do.call( rbind.data.frame, big_features )
path_big_features = './DATA/SUMM/big_features.csv'

fwrite(big_features, path_big_features)
```

```{r}
big_features = fread( path_big_features )
```


```{r}
#crossplot of average ordersize in prior and ordersize in train
colnames( big_features )

attach(big_features)
plot(ave_ordersize_in_prior, ordersize_in_train)
plot(lst_ordersize_in_prior, ordersize_in_train)

corr.plot(ave_ordersize_in_prior, ordersize_in_train, xlab ='ave_ordersize_in_prior', ylab = 'ordersize_in_train' )

corr.plot(lst_ordersize_in_prior, ordersize_in_train, xlab ='lst_ordersize_in_prior', ylab = 'ordersize_in_train' )  

corr.plot(ave_ordersize_in_prior, lst_ordersize_in_prior, xlab ='ave_ordersize_in_prior', ylab = 'lst_ordersize_in_prior' )  
detach(big_features)
```
```{r}
#crossplot of average proportion reordered in prior and proportion reordered in train
colnames( big_features )

attach(big_features)
plot(ave_prop_reordered_in_prior, prop_reordered_in_train)
plot(lst_prop_reordered_in_prior, prop_reordered_in_train)

corr.plot(ave_prop_reordered_in_prior, prop_reordered_in_train, xlab ='ave_prop_reordered_in_prior', ylab = 'prop_reordered_in_train' )

corr.plot(lst_prop_reordered_in_prior, prop_reordered_in_train, xlab ='lst_prop_reordered_in_prior', ylab = 'prop_reordered_in_train' )  

corr.plot(ave_prop_reordered_in_prior, lst_prop_reordered_in_prior, xlab ='ave_prop_reordered_in_prior', ylab = 'lst_prop_reordered_in_prior' )  
detach(big_features)
```
