---
title: "Instacart: Exploratory Analysis"
output: html_notebook
---

This part is based on the code provided at https://www.kaggle.com/philippsp/exploratory-analysis-instacart by Philipp Spachtholz. 

```{r "setup", include=FALSE}

cat("\014")  # clear console
rm(list=ls(all=TRUE))

require("knitr")
opts_knit$set(root.dir = 'c:/BBB/PROPACAD/CAPSTONE/BASKET/')
```


```{r}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)

orders <- fread('./DATA/orders.csv')
products <- fread('./DATA/products.csv')
order_products_train <- fread('./DATA/order_products__train.csv')
order_products_prior <- fread('./DATA/order_products__prior.csv')
aisles <- fread('./DATA/aisles.csv')
departments <- fread('./DATA/departments.csv')
```

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

```{r}
head(orders)
```

```{r}
head(products)
dim(products)
```


```{r}
head(order_products_train, n = 20)
dim( order_products_train )
```


```{r}
head(order_products_prior, n = 20)
dim(order_products_prior)
```

```{r}
head(aisles, n = 20)
```

```{r}
head(departments, n = 20)
```

Recode variables


```{r}
orders      <- orders      %>% mutate(order_hour_of_day = as.numeric(order_hour_of_day), eval_set = as.factor(eval_set))
products    <- products    %>% mutate(product_name      = as.factor(product_name))
aisles      <- aisles      %>% mutate(aisle             = as.factor(aisle))
departments <- departments %>% mutate(department        = as.factor(department))
```

```{r}
orders
```

Create for each customer a density matrix. For those who repeatedly order the same products this matrix will be dense, and for
those who order different products each time this matrix will be sparse.

```{r}
user_id_i = 5

user_id_i_prior <- orders %>% filter(user_id==user_id_i, eval_set == 'prior' ) 
user_id_i_prior <- user_id_i_prior %>% mutate( order_id_num = seq( nrow(user_id_i_prior) ) ) %>% 
                                       select( order_id,order_id_num, user_id )  %>% 
                                       left_join(order_products_prior,by="order_id") %>%
                                       left_join(products, by = 'product_id')
user_id_i_prior <- user_id_i_prior %>% mutate( order_id_num = as.factor(order_id_num), product_id = as.factor(product_id) )
user_id_i_prior

user_id_i_prior_num_order   = length( levels(user_id_i_prior$order_id_num  ) )
user_id_i_prior_num_product = length( levels(user_id_i_prior$product_id) )

#create an empty table
user_id_i_prior_tbl = matrix( 0, nrow = user_id_i_prior_num_product, ncol = user_id_i_prior_num_order)
rownames(user_id_i_prior_tbl) = unique( user_id_i_prior$product_name )
colnames(user_id_i_prior_tbl) = levels(user_id_i_prior$order_id_num  )

#insert values in the table
for( i in seq(user_id_i_prior_num_order) ){
  
  order_id_num_j = levels(user_id_i_prior$order_id_num  )[i]; order_id_num_j
  product_name   = user_id_i_prior %>% filter( order_id_num == order_id_num_j )
  product_name
  
  product_name   = product_name %>% select(product_name)
  product_name
  
  #convert from factor back to character
  product_name %>% mutate_if(is.factor, as.character) -> product_name
  product_name
  
  row_indx = as.vector( sapply( product_name[,1], function(x) which( x == rownames(user_id_i_prior_tbl) ) ) )
  row_indx
  
  user_id_i_prior_tbl[row_indx,order_id_num_j] = 1
}

print( paste( "number of orders in prior: "  , user_id_i_prior_num_order   ) )
print( paste( "number of products in prior: ", user_id_i_prior_num_product ) )



```

```{r}
user_id_i_train <- orders %>% filter(user_id==user_id_i, eval_set == 'train')  %>% 
                              select( order_id, user_id )  %>% 
                              left_join(order_products_train,by="order_id") %>%
                              left_join(products, by = 'product_id')
user_id_i_train
```


```{r}

```
